Compositional equivalences based on Open pNets

### 2020-04-24 2016-Weak-FH-Bisim/WeakBisim.tex
skype comments

We denote $y\gets e$ a substitution. The application of the substitution is denoted
$\subst{y\gets e}$, the operation replaces in a term all occurrences 
of the variable $y$ by the expression $e$ (note that the absence of binders makes this operation 
trivial). 
We use sets of assignments of the form $x_i:=e_i$, ranged over by $\Post$; $\subst{\Post}$ is the substitution that applies all the substitutions defined by $\Post$ in a parallel manner. $\otimes$ is the composition operator on substitutions, such that for any term $t$ we have: $t\subst{\Post\shortotimes\Post'} = (t\subst{\Post})\subst{\Post'}$.

@@

\begin{definition}[pNets]\label{def-pnets}
A pNet $\pNet$ is a hierarchical structure where leaves are pLTSs and holes:\\
$\pNet\triangleq \pLTS~|~\mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, \symb{SV}_k^{k\in 
K}\myrangle$
where
\begin{itemize}
\item[$\bullet$] $\pNet_i^{i\in I}$ is the family of sub-pNets indexed over $I$.
%  $\pNet_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which sub-pNets are indexed.

@@

\item The set of variables of a pNet $P$, denoted $vars(P)$ is the set of variables of  all pLTSs that compose $P$.

@@

\begin{definition}[Weak open transition]\label{def:weakOT}
A weak open transition over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
\begin{mathpar}
 \openrule
         {
           \gamma_j^{j\in J'},\Pred,\Post}
         {s \OTWeakarrow {\alpha} s'}
 \end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\gamma_j$
        is a list of transitions of the hole $j$, with each element of the list in $\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ and \Post\ are defined similarly to Definition~\ref{def:OT}. We use $\WT$ to range over sets of weak open transitions.

@@

\hl{Now focus on OTs}\footnote{In the appendix we use OT to abbreviate ``open transition'' and WOT to stand for ``weak open transition.''} of the subnets. For each $m\in I_1\uplus I_2$ we have one of the two 
following OT:\\
either $m$ in $I_1$
\[
\pNet_m \models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
or, $m$ in $I_2$
\[{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}}}\]

\hl{Only elements of} $(I_1\uplus I_2)\backslash\{j_0\}$ and  are useful to assert the premise for reduction of $P$; the last 
one ensures the open transition for the pNet $Q$ (note that $Q$ is at place $j_0$, and by 
definition of the open transition 
for $P[Q]_{j_0}$, 
$L_{j_0}=L_Q$, and $J_{j_0}=	J\cap\Holes(Q)$):\\[-2ex]
	\[Q\models{\openrule
		{
			\beta_j^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_{j_0}}
			\ostate{(s_i')^{\, i\in L_Q}}}
	}\]
\TODO{j'ai du mal, je reviens}

@@

\end{mathpar}
Moreover, by using Lemma \ref{lem-rel-OT-WOT} we know that:
\begin{mathpar}
    \openrule
         {
           \gamma_j^{j\in J'},\Pred_{OT},\Post_{OT}}
         {s \OTWeakarrow {\alpha} s'} \in \WT_1
         \Rightarrow
\bigg(\forall i \in [0..n].\openrule
    {
       \beta_{1ij}^{j\in J_1'},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T}_1  \wedge
\\{\qquad\qquad\qquad\qquad\qquad}
\openrule
         {
           \beta_{2j}^{j\in J_2'},\Pred_2,\Post_2 }
         {s_{20} \OTarrow {\alpha'} s_{21}} \in \mathcal{T}_1
 \wedge
\forall i \in [0..m].\openrule
         {
           \beta_{3ij}^{j\in J_3'},\Pred_{3i},  \Post_{3i}  }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}} \in \mathcal{T}_1
\bigg)
\end{mathpar}
where\\
$
\displaystyle{\alpha=\alpha'\subst{\bigotimes_{j=n}^{0}\Post_{1j}}}\\
s=s_{10} \wedge s_{1(n+1)}=s_{20} \wedge s_{21} = s_{30} \wedge s_{3(m+1)}=s'\\
\displaystyle{
\gamma_j^{j\in J'}=\mybigdotcup_{i=0}^n \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0} \Post_{1j} } }  \dotcup  \vis{\set{\beta_2}\subst{\bigotimes_{j=n}^{0}\Post_{1j}}} \dotcup}\\
{}~\qquad\qquad\!\displaystyle{\mybigdotcup_{i=0}^m \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}} }}
\\
\Pred_{OT}=\bigwedge_{i=0}^n\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=n}^{0}\Post_{1j}}\land\\ 
{}~\qquad\qquad\Big(\bigwedge_{i=0}^m\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}}\Big)\\
\Post_{OT}=\bigotimes_{j=m}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}$

>>>>

We denote $y\gets e$ a substitution. The application of the substitution is denoted
$\subst{y\gets e}$, the operation replaces in a term all occurrences 
of the variable $y$ by the expression $e$. 
$\Post$ ranges over (indexed) sets of substitutions; $\subst{\Post}$ is the substitution that applies all the substitutions defined by $\Post$ in a parallel manner. $\otimes$ is the composition operator on substitutions, such that for any term $t$ we have: $t\subst{\Post\shortotimes\Post'} = (t\subst{\Post'})\subst{\Post}$.

@@

\begin{definition}[pNets]\label{def-pnets}
A pNet $\pNet$ is a hierarchical structure where leaves are pLTSs and holes:\\
$\pNet\triangleq \pLTS~|~\mylangle \pNet_i^{i\in I}, \Sort_j^{j\in J}, \symb{SV}_k^{k\in 
K}\myrangle$
where
\begin{itemize}
\item[$\bullet$] $\pNet_i^{i\in I}$ is the family of sub-pNets indexed over $I$. $vars(P_i)$ and $vars(P_j)$ must be disjoint for $i\neq j$.
%  $\pNet_i^{i\in I}$ is a family of sub-pNets where $I\in\I_\P$ is the set over which sub-pNets are indexed.

@@

\item The set of variables of a pNet $P$, denoted $vars(P)$ is disjoint union the set of variables of  all pLTSs that compose $P$.

@@

\begin{definition}[Weak open transition (WOT)]\label{def:weakOT}
A weak open transition over a
	set $J$ of holes with sorts $\Sort_j^{j\in J}$ and a set of states $\mathcal{S}$ is 
	a structure of the form:	
\begin{mathpar}
 \openrule
         {
           \gamma_j^{j\in J'},\Pred,\Post}
         {s \OTWeakarrow {\alpha} s'}
 \end{mathpar}
	Where $J'\subseteq J$, $s, s'\in\mathcal{S}$ and $\gamma_j$
        is a list of transitions of the hole $j$, with each element of the list in $\Sort_j$. $\alpha$ is an action 
        label denoting the resulting action
        of this open transition. \Pred\ and \Post\ are defined similarly to Definition~\ref{def:OT}. We use $\WT$ to range over sets of weak open transitions.

@@

Now focus on the OTs of the subnets. For each $m\in I_1\uplus I_2$ we have one of the two 
following OT:\\
either $m$ in $I_1$
\[
\pNet_m \models\openrule
    	{
    	\beta_{j}^{j\in J_m}, \Pred_m, \Post_m}
    	{\ostate{s_{i}^{i \in L_m}} \OTarrow {\alpha_m}
    		\ostate{(s_i^\prime)^{i\in L_m}}}\]
or, $m$ in $I_2$
\[{ \pNet_m 
    	 \models
    	\openrule
    	{\emptyset, \Pred_m, \Post_m}
    	{\ostate{s_m} \OTarrow {\alpha_m}
    		\ostate{s_m'}}}\]

Only elements of $(I_1\uplus I_2)\backslash\{j_0\}$   are useful to assert the premise for reduction of $P$; the last 
one ensures the open transition for the pNet $Q$ (note that $Q$ is at place $j_0$, and by 
definition of the open transition 
for $P[Q]_{j_0}$, 
$L_{j_0}=L_Q$, and $J_{j_0}=	J\cap\Holes(Q)$):\\[-2ex]
	\[Q\models{\openrule
		{
			\beta_j^{j\in J\cap\Holes(Q)}, \Pred_{j_0},  
			\Post\,''}
		{\ostate{s_i^{i\in L_Q}} \OTarrow {\alpha_{j_0}}
			\ostate{(s_i')^{\, i\in L_Q}}}
	}\]
\TODO{j'ai du mal, je reviens}

@@

\end{mathpar}
Moreover, by using Lemma \ref{lem-rel-OT-WOT} we know that:
\begin{mathpar}
    \openrule
         {
           \gamma_j^{j\in J'},\Pred_{OT},\Post_{OT}}
         {s \OTWeakarrow {\alpha} s'} \in \WT_1
         \Rightarrow
\bigg(\forall i \in [0..n].\openrule
    {
       \beta_{1ij}^{j\in J_1'},\Pred_{1i},\Post_{1i}   }
         {s_{1i} \OTarrow {\tau} s_{1(i+1)}} \in \mathcal{T}_1  \wedge
\\{\qquad\qquad\qquad\qquad\qquad}
\openrule
         {
           \beta_{2j}^{j\in J_2'},\Pred_2,\Post_2 }
         {s_{20} \OTarrow {\alpha'} s_{21}} \in \mathcal{T}_1
 \wedge
\forall i \in [0..m].\openrule
         {
           \beta_{3ij}^{j\in J_3'},\Pred_{3i},  \Post_{3i}  }
         {s_{3i} \OTarrow {\tau} s_{3(i+1)}} \in \mathcal{T}_1
\bigg)
\end{mathpar}
where
\begin{align*}
s=s_{10} &\wedge s_{1(n+1)}=s_{20} \wedge s_{21} = s_{30} \wedge s_{3(m+1)}=s'\\
\alpha=&\alpha'\subst{\bigotimes_{j=n}^{0}\Post_{1j}}\\
\gamma_j^{j\in J'}=&\mybigdotcup_{i=0}^n \vis{\set{\beta_{1i}}\subst{\bigotimes_{j=i-1}^{0} \Post_{1j} } }  \dotcup  \vis{\set{\beta_2}\subst{\bigotimes_{j=n}^{0}\Post_{1j}}} \dotcup\\
&\mybigdotcup_{i=0}^m \vis{\set{\beta_{3i}} \subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}} }
\\
\Pred_{OT}=&\bigwedge_{i=0}^n\Pred_{1i}\subst{\bigotimes_{j=i-1}^{0}\Post_{1j}}\land\Pred_2 \subst{\bigotimes_{j=n}^{0}\Post_{1j}}\land\\ 
&\qquad\Big(\bigwedge_{i=0}^m\Pred_{3i}\subst{\bigotimes_{j=i-1}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}}\Big)\\
\Post_{OT}=&\bigotimes_{j=m}^{0}\Post_{3j}\shortotimes\Post_2\shortotimes\bigotimes_{j=n}^{0}\Post_{1j}
\end{align*}
